Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2014-12-17 ms.tex
fixed mccourt mania

\citet{Ho:2009a} calculates the accretion rates onto a sample of
early-type galaxies both by using X-ray observations to determine the
Bondi accretion rate, and by estimating the contributions from the
mass loss of evolved stars.  Both methods lead him to conclude that
the available reservoir of gas is more than sufficient to power
low-luminosity AGN, assuming the s
If cooling becomes important in the flow ($v_{w} < v_{\rm crit}$), it may lead to thermal instabilities.  There are two kinds of ``instabilities", global and local.  The global instability is simply that, if cooling becomes very rapid, the gas loses thermal pressure and starts to inflow faster, i.e. the entire CNM accretes.  In
practice this should could manifest as a large increase in the stagnation radius, due an effective {\it decrease} in the effective heating rate.  This global ``instability" is not a true instability insofar as the gas does not necessarily clump locally into low-T clouds.

However, \citet{McCourt+12} argue that even if cooling is exactly balanced by heating everywhere, there is still a local thermal instability that operates to turn hot gas into cold clouds.  However, this local instability only causes a significant local drop-out of mass from the hot medium (to low-T clouds), if the cooling timescale,
\begin{align}
\frac{\tcool}{\tff} &\equiv \frac{3\rho k T/2 \mu \mp}{\dot{q}_{\rm cool}} \approx \frac{9}{5} \left. \frac{q \vw^2/2}{\dot{q}_{\rm cool}}\right|_{\rs}
\end{align} 
is locally shorter than the approximately 10 times the free-fall
timescale $t_{\rm ff}$, where the last equality is evaluated at the stagnation radius and makes use of equations~(\ref{eq:Tanalytic}) and (\ref{eq:rhors2}).  We thus see that the condition $v_w \gtrsim v_{\rm crit}$ is also the condition for local instability, although the prefactor may be slightly different.
tandard radiative efficiency $\eta
\sim 0.1$ for thin disk accretion.  Indeed, there is now strong
observational evidence that the low levels of AGN activity in
quiescent galaxies results from accretion proceeding in a radiatively
inefficient mode (\citealt{Yuan&Narayan14}), either due to the
gravitationally-released energy being advected across the SMBH horizon
(e.g.~\citealt{Narayan&Yi95}) or due to only a fraction of the gas
inflowing on large scales ultimately reaching the SMBH
(e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).

@@

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:sample}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe our
results.  In $\S\ref{sec:applications}$ we describe applicaiton of our
results, including a comparison of our gas profiles to those inferred
from {\it Chandra} observations and to the measured $L_X-\Mbh$
relationship.  In $\S\ref{sec:discussion}$ we discuss our results.  In
$\S\ref{sec:summary}$ we summarize our results.
\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\citet{LauerFaber+:2007a} use Hubble WFPC2 imaging to measure the
radial surface brightness profiles for hundreds of nearby early type galaxies. The measured surface brightness profile is generally well-fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b}.
\end{equation}

The Nuker profile is a broken power law which transitions from an inner power law slope, $\Gamma$, to an outer power law slope, $\beta$, at a break radius, $\rb$.  Assuming spherical symmetry, this corresponds to a stellar density $\rhostar \sim r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\sim r^{-1-\beta}$ for $r \gg \rb$, thus motivating us to adopt a broken power-law profile
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star(\rsoi) \left(r/\rsoi\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star(r_b) \left(r/r_b\right)^{-1-\beta} & r > r_b,
\label{eq:rhostar}
\end{cases}
\end{align}
where $\rho_\star(\rsoi)$ is the stellar density at the radius of the black hole sphere of influence, 
\be
\rsoi \simeq G \Mbh/\sigma^2 \approx 14 M_{\bullet,8}^{0.6}\,{\rm pc},
\label{eq:rsoi}
\ee
and the final equality in (\ref{eq:rsoi}) employs a recent version of the $\Mbh-\sigma$ relationship (\citealt{Gultekin+09}),
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}
Although the use of equation (\ref{eq:Msigma}) is questionable for low mass black holes (e.g., \citealt{Greene&Ho07}), our results are not overly sensitive to this assumption.

>>>>

\citet{Ho:2009a} calculates the accretion rates onto a sample of
early-type galaxies both by using X-ray observations to determine the
Bondi accretion rate, and by estimating the contributions from the
mass loss of evolved stars.  Both methods lead him to conclude that
the available reservoir of gas is more than sufficient to power
low-luminosity AGN, assuming the standard radiative efficiency $\eta
\sim 0.1$ for thin disk accretion.  Indeed, there is now strong
observational evidence that the low levels of AGN activity in
quiescent galaxies results from accretion proceeding in a radiatively
inefficient mode (\citealt{Yuan&Narayan14}), either due to the
gravitationally-released energy being advected across the SMBH horizon
(e.g.~\citealt{Narayan&Yi95}) or due to only a fraction of the gas
inflowing on large scales ultimately reaching the SMBH
(e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).

@@

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:sample}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe our
results.  In $\S\ref{sec:applications}$ we describe applicaiton of our
results, including a comparison of our gas profiles to those inferred
from {\it Chandra} observations and to the measured $L_X-\Mbh$
relationship.  In $\S\ref{sec:discussion}$ we discuss our results.  In
$\S\ref{sec:summary}$ we summarize our results.
\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\citet{LauerFaber+:2007a} use Hubble WFPC2 imaging to measure the
radial surface brightness profiles for hundreds of nearby early type galaxies. The measured surface brightness profile is generally well-fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b}.
\end{equation}
The Nuker profile is a broken power law which transitions from an inner power law slope, $\Gamma$, to an outer power law slope, $\beta$, at a break radius, $\rb$.  Assuming spherical symmetry, this corresponds to a stellar density $\rhostar \sim r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\sim r^{-1-\beta}$ for $r \gg \rb$, thus motivating us to adopt a broken power-law profile
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star(\rsoi) \left(r/\rsoi\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star(r_b) \left(r/r_b\right)^{-1-\beta} & r > r_b,
\label{eq:rhostar}
\end{cases}
\end{align}
where $\rho_\star(\rsoi)$ is the stellar density at the radius of the black hole sphere of influence, 
\be
\rsoi \simeq G \Mbh/\sigma^2 \approx 14 M_{\bullet,8}^{0.6}\,{\rm pc},
\label{eq:rsoi}
\ee
and the final equality in (\ref{eq:rsoi}) employs a recent version of the $\Mbh-\sigma$ relationship (\citealt{Gultekin+09}),
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}
Although the use of equation (\ref{eq:Msigma}) is questionable for low mass black holes (e.g., \citealt{Greene&Ho07}), our results are not overly sensitive to this assumption.
